Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-07 edubots_surveys.Rnw

Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning
experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

@@

In this study, data collection involved two phases:
\begin{itemize}
\item The first phase aimed to collect feedback from students about their actual
  information about students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who do they like to be
  with in the class messaging groups, and their expectations of the chatbots in
  terms of assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  applications, such as Telegram or WhatsApp as they are familiar to
  them. Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for course's resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. This is what is going to be presented in this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed\textcolor{black}{; this will be explained more fully in the methodology section}. The responses' analysis
entailed the creation of a second survey to which 200 teachers' responses were
collected\textcolor{black}{, with the methodology and collective characteristics that will be explained below. The questions of the second survey focused on discussing teachers'
needs, opinions and preferences of the development of future technology-enhanced
tools and its potential impact on educational institutions’ policies. In order to assess these, this paper investigated the following research questions}:

\begin{itemize}
\item RQ1 - Do teachers use now or want to use messaging apps in their classes?
\item RQ2 - Which chatbots' features would teachers find useful for this purpose?
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these questions through the analysis of questions and
answers in the two aforementioned surveys. Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.

The remainder of the paper is organized as follows: first, an overview of what
current research has found about the use of messaging applications, including
chatbots, in the classroom is described. The methodology used in the surveys is presented in
Section \ref{sec:meth}, and the results of the surveys are presented next in
Section \ref{sec:res}. Finally, we discuss these results and conclude with a
series of recommendations in the section that closes the paper.

>>>>

Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning
experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot will allow teachers to spend more time on  other topics that students struggle with.

@@

In this study, data collection involved two phases:
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their
  students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who they like to be
  with in the class messaging groups, and their expectations of chatbots 
  in assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  familiar applications, such as Telegram or WhatsApp. 
  Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed\textcolor{black}{; this will be detailed in Section \ref{sec:meth}}. The responses' analysis
led to the preparation of a second survey for teachers and 200 responses were
collected\textcolor{black}{, again following the methodology that will be explained later. The questions of the second survey focused on asking about teachers'
needs, opinions and preferences of the development of future technology-enhanced
tools and its potential impact on educational institutions policies. In order to assess these, this paper layed out the following research questions}:

\begin{itemize}
\item RQ1 - Do teachers use now or want to use messaging apps in their classes?
\item RQ2 - Which chatbots' features would teachers find useful for this purpose?
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - Which kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these research questions through the analysis of the two aforementioned surveys. 
Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.

The remainder of the paper is organized as follows: first, an overview of what
current research has found about the use of messaging applications, including
chatbots, in the classroom is described. The methodology used in the surveys is presented in
Section \ref{sec:meth}, and the results of the surveys are presented next in
Section \ref{sec:res}. Finally, we discuss these results and conclude with a
series of recommendations for the successful deployment of chatbots in higher education.
